---
title: "TidyVerseAssignment"
author: "Jonnathan Zuna Largo"
format: html
editor: visual
---

## Data Science Careers & Salaries 2025

Source: <https://www.kaggle.com/datasets/nalisha/data-science-careers-and-salaries-2025>.

## Code

```{r}
library(tidyverse)

url <- "https://raw.githubusercontent.com/JZunaRepo/TidyVerseAssignment/refs/heads/main/data_science_job_posts_2025.csv"

jobs <- read.csv(
  file = url
)
```

```{r}
glimpse(jobs)
```

Given that the salary has a range and values that contain quotations, we need to strip these signs and commas and extract the numbers to be able to compute them

```{r}
jobs <- jobs %>%
  mutate(
    salary_clean = str_remove_all(salary, "[€,]"),
    salary_low = as.numeric(str_extract(salary_clean, "^\\d+")),
    salary_high = as.numeric(str_extract(salary_clean, "\\d+$")),
    salary_mid = if_else(salary_low != salary_high,
                         (salary_low + salary_high) / 2,
                         salary_low)
  )
```

```{r}
jobs %>%
  select(salary, salary_low, salary_high, salary_mid) %>%
  head(10)
```

The skills column also need cleaning. We can achieve this by formatting and expand each skill to its own row

```{r}
skills_long <- jobs %>%
  select(job_title, seniority_level, industry, skills) %>%
  mutate(
    skills_clean = str_remove_all(skills, "[\\[\\]\\']"),
    skills_list = str_split(skills_clean, ",\\s*")
  ) %>%
  unnest(skills_list) %>%
  mutate(
    skills_list = str_trim(skills_list)
  ) %>%
  filter(skills_list != "")
head(skills_long, 10)
```

Now we can observe all the skills in order by most frequent skill

```{r}
top_skills <- skills_long %>%
  count(skills_list, sort = TRUE) %>%
  slice_head(n = 15)

top_skills
```

We can also graph and visualize the skills by frequency

```{r}
top_skills %>%
  mutate(skills_list = fct_reorder(skills_list, n)) %>%
  ggplot(aes(x = n, y = skills_list)) +
  geom_col(fill = "#2c7bb6") +
  geom_text(aes(label = n), hjust = -0.2, size = 3.5) +
  labs(
    title = "Top 15 Most In-Demand Skills in Data Science Job Postings (2025)",
    x = "Number of Job Postings",
    y = "Skill",
    caption = "Source: Data Science Job Posts 2025"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, face = "bold"),
    axis.text.y = element_text(size = 11)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)))
```

We can also observe the salary distributrion by seniority level

```{r}
jobs %>%
  filter(seniority_level != "") %>%
  ggplot(aes(x = fct_reorder(seniority_level, salary_mid), y = salary_mid)) +
  geom_boxplot(fill = "#2c7bb6", alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.25, color = "#1a1a2e", size = 1.2) +
  coord_cartesian(ylim = c(0, 400000)) +
  labs(
    title = "Salary Distribution by Seniority Level (2025)",
    x = "Seniority Level",
    y = "Salary (€)",
    caption = "Source: Data Science Job Posts 2025"
  ) +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, face = "bold"),
    axis.text.x = element_text(size = 11)
  )
```

And look at top companies hiring data scientists roles

```{r}
jobs %>%
  filter(industry != "") %>%
  count(industry, sort = TRUE) %>%
  slice_head(n = 10) %>%
  mutate(industry = fct_reorder(industry, n)) %>%
  ggplot(aes(x = n, y = industry)) +
  geom_col(fill = "#2c7bb6") +
  geom_text(aes(label = n), hjust = -0.2, size = 3.5) +
  labs(
    title = "Top 10 Industries Hiring Data Scientists (2025)",
    x = "Number of Job Postings",
    y = "Industry",
    caption = "Source: Data Science Job Posts 2025"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 13, face = "bold"),
    axis.text.y = element_text(size = 11)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)))
```

## Conclusion

This bignette wakled trhough a complete data wrangling and visualization workflow using the following pakacges from Tidyverse: dplyr, tidyr, and ggplot2. These packages were applied to the kaggle datasets of job postings from 2025. The analysis revealed that Python and machine learning are by far the most in-demand skills in the current job market, with SQL and R close behind. Technology unsurprisingly leads all industries in both hiring volume and average compensation, though Finance and Healthcare also showed strong representation.
